---
title: "Merge and Imputation w2 and w3"
author: "Bharvi Dhall"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=5, figs="figsBD/")
```

Goal : Take school type values from wave 3 as it is and impute the missing ones from wave 2.


## Loading Libraries

```{r Loading_Libraries}

library(tidyverse)
library(here)
library(labelled)
library(haven)

```

## Read the Robject from wave 3
Using only variables shortlisted for GUI conference.

data from wave 3 -> dw3
Unique Identifier Household ID -> ID
No of boys enrolled -> p3q4a
No of girls enrolled -> p3q4b


```{r}

here::i_am("CATAv2/00b_TD_datasetup.Rmd")
wave3 <- readRDS(here("CATAv2","R_ob","Cata_data_W3_sub.RDS"))
dim(wave3)
# [1] 6216   33

#View(wave3)



#Looking at the unique Identifier Household id
length(unique(wave3$ID))
#6216

#View(wave3[1:10, c("ID","p3q4a","p3q4b")])



```

## Principal's Questionnaire: No of boys and girls wave 3
Note: taking don't know as other (which is non zero)
other_level (an argument of fct_collapse : Value of level used for "other" values. Always placed at end of levels.)


```{r}

dw3<-wave3
#dw3 %>% count(across(p3q4b:p3q4a)) %>% View()

dw3 %>% mutate(p3q4b = fct_collapse(p3q4b, no = "0" , NULL = NA, other_level = "yes")) -> dw3
dw3 %>% mutate(p3q4a = fct_collapse(p3q4a, no = "0" , NULL = NA, other_level = "yes")) -> dw3



# create a column school_type which indicated whether school is single-sex or mixed

dw3 <- dw3 %>% mutate(school_type= case_when(
   p3q4b == "yes" & p3q4a == "yes" ~ "mixed",
   p3q4b == "yes" | p3q4a == "yes" ~ "single"))




```


## Read the Robject from wave 2

data from wave 2 -> dw2
Unique Identifier Household ID -> zid01

```{r}

dw2 <- readRDS(here("CATAv2","R_ob","Cata_data_W2.RDS"))
dim(dw2)
# [1] 7525  913

#looking at the unique identifier Household ID(zid01)
length(unique(dw2$zid01))
#7525



```

## Get unique ID and no of boys and girls from wave 2

No of boys enrolled -> bhpc04a
No of girls enrolled -> bhpc04b

```{r}

temp.dw2 <- dw2 %>% 
  select(zid01, bhpc04a, bhpc04b)


```

## Principal's Questionnaire: No of boys and girls wave 2


```{r}

#temp.dw2 %>% count(across(bhpc04a:bhpc04b)) %>% View()

# for boys
temp.dw2 %>% mutate(bhpc04a = fct_collapse(bhpc04a, no = "0" , NULL = NA, other_level = "yes")) -> temp.dw2

# for girls
temp.dw2 %>% mutate(bhpc04b = fct_collapse(bhpc04b, no = "0" , NULL = NA, other_level = "yes")) -> temp.dw2


# create a column school_type which indicated whether school is single-sex or mixed

temp.dw2 <- temp.dw2 %>% mutate(school_type_w2= case_when(
   bhpc04a == "yes" & bhpc04b == "yes" ~ "mixed",
   bhpc04a == "yes" | bhpc04b == "yes" ~ "single"))

# only taking school_type_w2
temp2 <- temp.dw2 %>% 
  select(zid01,school_type_w2)

```

Note: wave 2 had only 400 approx missing whereas wave 3 had over 1300 missing for these columns

## Merge w2 and w3


```{r}

w3_sub <- merge(dw3,temp2,by.x = "ID", by.y = "zid01", all.x=TRUE)

# Na's wave 3
sum(is.na(w3_sub$school_type)) #1359

# Na's wave 2
sum(is.na(w3_sub$school_type_w2))  #405


```

Is the school type consistent across the two waves

```{r}

#looking at the complete cases

temp_df <- w3_sub[,c("ID","school_type","school_type_w2")]
temp_df <- temp_df[complete.cases(temp_df),]

#View(temp_df)
#dim(temp_df) #4548

#match columns
all(temp_df$school_type ==  temp_df$school_type_w2)
#False

#looking at count of matched
sum(temp_df$school_type ==  temp_df$school_type_w2) #4249

#Mismatched
sum(temp_df$school_type != temp_df$school_type_w2)  #299

temp_df %>% 
  filter(temp_df$school_type != temp_df$school_type_w2) 

temp_mismatch <- temp_df %>% 
  filter(temp_df$school_type != temp_df$school_type_w2) %>%
  select(ID)


# Wave 3
wave3[wave3$ID %in% unlist(temp_mismatch),c("p3q4a","p3q4b")]
# Wave 2
dw2[dw2$zid01 %in% unlist(temp_mismatch),c("bhpc04a","bhpc04b")]


#View(w3_sub)


```

Growing Up in Ireland is a longitudinal study based on a fixed panel
design. This means that the project follows the children and their families who were recruited into the study at 9 years of age for re-interview on several subsequent occasions. 

We can see that the school type is not consistent in the two waves. This can happen when the study child has changed schools.

Now we will keep the school type obtained from wave 3 as it is and impute the missing one's from wave 2.

```{r}

w3_sub <- w3_sub %>% 
    mutate(imp_school_type = coalesce(school_type,school_type_w2)) 


sum(is.na(w3_sub$imp_school_type))  #96

sum(is.na(w3_sub$school_type)) # 1359 were missing
sum(w3_sub$school_type ==  w3_sub$imp_school_type, na.rm = TRUE)

```

Adding imp_school_type to wave 3 data
```{r}
wave3$imp_school_type <- w3_sub$imp_school_type
df_names <- wave3 %>%sjlabelled::get_label() %>% enframe()
#View(df_names)

```

## Looking at the pricipal's questionnaire to get the school ID


```{r}

w3 <- readRDS(here('Data', 'RObjects_created_using_GUI','Wave3_GUI_17yr.RDS'))

df3 <- w3 %>%
  mutate_if(
    is.labelled,
    ~as_factor(.x)
  )

```


```{r}
# WAVE 3
# Looking at the responses from principles to see if we can figure out unique schools
teach3 <- select(df3, c(ID, w3prin,starts_with("p3q"))) 
teach3 %>% count(w3prin) # 1359 NA

dim(teach3)

uteach3 <- distinct(select(teach3, -ID))
dim(uteach3) 
# [1] 463 145

uteach3$schoolID <- paste0("W3", 1:nrow(uteach3))

# Left join principal ID with wave3
teach <- left_join(teach3, uteach3)
dim(teach)

prin3 <- teach %>% select(ID, schoolID)

wave3_sub <- left_join(wave3,prin3,by = "ID")

wave3_sub %>% count(schoolID) 

```




## Merging Region with wave 1
Some Notes:

Variable Region: Urban or Rural refers to the location of the household and has been derived from the Primary Caregivers Questionnaire (Wave 1)- This variable is only available in Wave 1

Read R object from wave 1
```{r}

dw1 <- readRDS(here("CATAv2","R_ob","Cata_data_W1.RDS"))
dim(dw1)


#looking at the unique identifier Household ID(zid01)
length(unique(dw1$zid01))

#Looking at the region
levels(dw1$adsd60)

#subsetting wave 1 dataset
temp.dw1 <- dw1 %>%
  select(zid01,adsd60)

#merge wave 3 and wave 1 by left join

wave3_sub2 <- merge(wave3_sub,temp.dw1 ,by.x = "ID", by.y = "zid01", all.x=TRUE)

dim(wave3_sub2)
#[1] 6216   43

```




Saving as R object

```{r}

# Saving wave1_gui as .Rdata
saveRDS(wave3_sub2, here("CATAv2","R_ob", "Cata_w3_imputed.RDS" ))

```


