---
title: "Data Wrangling"
author: "Bharvi Dhall & Catherine Hurley"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=5, figs="figsBD/")
```
# Description
This file used the CATA Wave3 subset data object and performs data wrangling on it.

# Set file path

```{r}
library(here)
here()
  
```

# Loading Libraries

```{r Loading_Libraries}
library(tidyverse)
library(ggplot2)
library(forcats)
library(naniar)
library(sjlabelled)

```



# Read dataset 
Subset of GUI data set.
The columns are 11 career choices, other variables include gender, maths grades, parents education and expectation from parents

# Identifiers
ID- student ID
schoolID - schoolID

List of new variables
#parent 1
f14_were_you_born_in_ireland
c3_how_far_do_you_expect_yp_will_go_in_education_or_training

#parent 2
e14_were_you_born_in_ireland
b1_how_far_do_you_expect_yp_will_go_in_education_or_training

#young person citizen of ireland
k7a_are_you_a_citizen_of_ireland

# How often parents ask about- HOPAA
how_often_have_did_your_parent_asked_how_you_are_were_coping_with_the_amount_of_work_course_work_etc_for_your_courses

how_often_have_did_your_parents_discussed_how_you_are_getting_on_with_different_subjects_at_school_college

how_often_have_did_your_parent_discussed_your_plans_for_the_future

how_often_have_did_your_parent_asked_how_you_are_were_getting_on_with_your_teachers_lecturers

how_often_have_did_your_parent_discussed_how_you_did_in_tests_or_exams

c13a_agree_or_not_i_dont_need_to_have_a_university_institute_of_technology_degree_to_get_the_kind_of_job_i_want_to_do

b17a_in_this_final_school_year_have_had_any_grinds_or_private_tuition_in_any_of_school_subjects_excluding_special_educational_needs_support

c3_what_do_you_think_you_are_most_likely_to_do_when_you_leave_school


# school

q5_describe_ethos_of_school

q4a_number_of_boys

q4b_number_of_girls

school_type_w2 represents whether school is single sex or mixed


## Read the data

```{r Reading_data, fig.height= 8, fig.width=12, message= FALSE, warnings = FALSE}

here::i_am("CATAv2/01_TD_DataWrangling.Rmd")
data  <- readRDS(here("CATAv2","R_ob", "Cata_w3_imputed.RDS"))
dim(data)
# [1] 6216   43

# Looking at the different variables present
#names(data)
#View(data)


# Getting long names for the columns next to short names
df_names <- data %>%sjlabelled::get_label() %>% enframe() %>% View()

#To view the data dictionary
#View(df_names)

#df_names <- df_names %>% mutate(short_name = c("income","training","interesting" , "flex_hrs","time_off","career","boss","help_soc","security","travel","other","gender","sclass","m_ed","f_ed","level_jcmaths","level_lcmaths","grade_lcmaths" ,"cog_naming","cog_maths","cog_vocab","p1.irish","p2.irish","yp.irish","p1.expe
Hi ct","p2.expect","subjects","work","teachers","future","exams", "degree_needed","post_school","suitable_accomodation" ,"boys", "girls","ethos","grinds" ))

# Renaming variables with shorter names
#names(data) <- df_names$short_name




df <- droplevels(data)


#View(data)
```

We are focusing on the question:
Which choices would you consider while choosing a job?

This question has 11 responses. Out of which 17 year olds have been asked to pick 3 that apply.

Each response has been coded as a variable with levels Yes and NA. Yes when students pick it, NA when they don't.

For the purpose of this analysis, all the NA's have been coded as No.

```{r}
df <- df %>%  mutate(across(starts_with("cq3g2"), ~as.numeric(.x)), across(starts_with("cq3g2"), ~replace_na(.x, 0)))

```


## looking at the NA values
```{r out.width="100%", message= FALSE, warnings = FALSE}

#total Na values
sum(is.na(df))


vis_miss(df[,-(1:11)])
# High amount of missing (over 30%) for f_ed,p2.irish, p2.expect, degree (93%) and level_lcmaths (88.75%)


```

# Data Cleaning and Manipulation

## Level of Education for Primary- pc3f1educ and Secondary Caregiver- sc3e1educ

```{r}

df %>% dplyr::level(pc3f1educ)



#levels(df$pc3f1educ)

df <- 
  df %>% mutate(pc3f1educ = fct_recode(pc3f1educ, NULL = "Don't Know",
                                       NULL ="Refusal",
                                  Prim = "None/Primary school", 
                                  JC = "Lower Secondary", 
                                  LC = "Hi Sec/TechVoc/UppSec+Tech/Voc", 
                                  No_degree = "Non Degree",
                                  Deg = "Primary Degree",
                                  Postgrad = "Postgrad"))
  

```

## gender- p2sexW3
```{r}
df %>% dplyr::count(p2sexW3)
```

## Cognitive maths (cog_maths)- CognitiveMathsTotal
```{r}

df %>% count(CognitiveMathsTotal, sort = TRUE)

df %>%
   mutate(CognitiveMathsTotal = as.numeric(CognitiveMathsTotal)) -> df
```

## Cognitive Naming (cog_naming)- CognitiveNamingTotal
```{r}

df %>% count(CognitiveNamingTotal, sort = TRUE)

df %>%
   mutate(CognitiveNamingTotal = as.numeric(CognitiveNamingTotal)) -> df
```

## cog_vocab- CognitiveVocabularyTotal
```{r}

df %>% count(CognitiveVocabularyTotal, sort = TRUE)

df %>%
   mutate(CognitiveVocabularyTotal = as.numeric(CognitiveVocabularyTotal)) -> df
```


## Combine all the cognitive scores to create a new variable by averaging them

```{r}

df %>%
  mutate(Cog_scores = rowMeans(dplyr::across(CognitiveNamingTotal:CognitiveVocabularyTotal))) -> df


df$CognitiveMathsTotal <- NULL
df$CognitiveNamingTotal <- NULL
df$CognitiveVocabularyTotal <- NULL
```

## YP TIPI conscientious Subscale W3

```{r}

df %>% count(w3cq_conscientious)

```

## hopa work subjects future teachers exams


```{r}

df <- 
  df %>% mutate(dplyr::across(cq3e1a:cq3e1f, ~ fct_recode(.x, "0" = "Don't Know",
                                      "0" = "Refusal",
                                    "1" = "Never or hardly ever",
                                    "1" ="A few times a year",
                                    "2" = "About once a month",
                                    "3" = "Several times a month",
                                    "4" = "Several times a week"))) %>%
  mutate(dplyr::across(cq3e1a:cq3e1f, as.numeric)) %>%
  mutate(hopa = rowMeans(dplyr::across(cq3e1a:cq3e1f))) %>%
  select(-(cq3e1a:cq3e1f))



```

## In thinking about career did you attend talks and take guidance cq3c2a:cq3c2e

```{r}

df <-
  df %>% mutate(dplyr::across(cq3c2a:cq3c2e, ~ fct_recode(.x, 
                                    "0" = "No",
                                    "1" = "Yes"))) %>%
  mutate(dplyr::across(cq3c2a:cq3c2e, ~abs(as.numeric(.x)-2))) %>%
  mutate(guidance = rowMeans(dplyr::across(cq3c2a:cq3c2e))) %>%
  select(-(cq3c2a:cq3c2e))

#View(df)
#Values 0 to 1. 0- no support at all




```

## p1empW3 - PC employment status wave 3
```{r}

df %>% count(p1empW3)

df <- 
  df %>% mutate(p1empW3 = fct_recode(p1empW3,
                                     Education = "At school/education",
                                     Work= "Work or Education",
                                    Unemployed = "Unemployed",
                                    Home = "Home duties / retired",
                                    Other = "Other"))

```

## w3equivinc - house income
```{r}

df %>% count(w3equivinc, sort = TRUE)

df %>%
   mutate(w3equivinc = as.numeric(w3equivinc)) -> df
```
## cq3g4f- How important are proffession and work

```{r}

 
df %>% count(cq3g4f)

df %>%
   mutate(cq3g4f = as.numeric(cq3g4f)) -> df

```

## Imp of family- cq3g4h

```{r}

df %>% count(cq3g4h)


df %>%
   mutate(cq3g4h = as.numeric(cq3g4h)) -> df


```

## cq3g4j - religion

```{r}

df %>% count(cq3g4j)


df %>%
   mutate(cq3g4j = as.numeric(cq3g4j)) -> df


```

## Did you take TY?

```{r}
df %>% count(cq3b4a)


df <- 
  df %>% mutate(cq3b4a = fct_recode(cq3b4a, NULL = "Don't Know",
                                    NULL = "Refusal",
                                    "Yes" = "Yes",
                                     "No"= "No"))

#levels(df$cq3b4a)
#View(df)
dim(df)
#[1] 6216   25

```

## School_type
```{r}
df %>% count(imp_school_type)
```

## No of boys - p3q4a and No of girls- p3q4b
We don't need these variables as we already have school type
```{r}

 #df %>% count(across(p3q4b:p3q4a)) %>% View()



df$p3q4b <- NULL
df$p3q4a <- NULL


```

## cq3a3ayr - when did the student left school? (can correspond to whether student left school or not)
This is a categorical variable with four levels: "Prior to 2015", "2015-2016", "Don't Know" and "NA".
NAs can be coded as "No" and other levels can be coded as "Yes".
```{r}
df %>% count(cq3a3ayr)

df <- df %>% mutate(cq3a3ayr=as.factor(replace_na(as.character(cq3a3ayr),"No")))

df <- 
  df %>% mutate(cq3a3ayr = fct_recode(cq3a3ayr, "Yes" = "2015-2016",
                                    "Yes" = "Don't Know",
                                    "No" = "No",
                                    "Yes"= "Prior to 2015"))
df %>% count(cq3a3ayr)

```

## cq3f1 How many hours per week on average in this job during term-time

```{r}

df %>% count(cq3f1)

df <- 
  df %>% mutate(cq3f1 = fct_recode(cq3f1, NULL = "Don't Know",
                                    "Yes" = "Yes",
                                     "No"= "No"))

```

## cq3f5 Do you ever do any work for a business owned or run by a member of your      family (paid or unpaid work)

```{r}

df %>% count(cq3f5)

df <- 
  df %>% mutate(cq3f5 = fct_recode(cq3f5, NULL = "Don't Know",
                                    "Yes" = "Yes",
                                     "No"= "No"))
```


## cq3f6 During this/your final school year had a short term work experience      placement, as part of your school curriculum

```{r}

df %>% count(cq3f6)

df <- 
  df %>% mutate(cq3f6 = fct_recode(cq3f6, NULL = "Don't Know",
                                    "Yes" = "Yes",
                                     "No"= "No"))

```


## region

```{r}

df %>% count(adsd60)

```



## Final dimensions of dataset

```{r}
dim(df)


```
31 vars left.



## Save rds
```{r out.width="100%", message= FALSE, warnings = FALSE}


#View(df)

saveRDS(df, here("CATAv2","R_ob","Cata_W3_sub_tidy.RDS" ))
```

