---
title: "02_TD_Graphs_for_Paper"
author: "Bharvi Dhall"
date: "`r Sys.Date()`"
output: html_document
---
Note- Run 00a_TD, 00b_TD, 01_TD before running this file
Install package performance to complete kniting file.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=10, fig.height=5, dpi = 300, fig.path = "GUIVIS_Plots/", warning = FALSE,message = FALSE)
```


This file  contains graphs for the paper


## Loading Library

```{r loading_libraries}
library(here)
library(sjlabelled) 
library(tidyverse)
library(ggplot2)
library(Hmisc) #for assigning labels
library(manyModelViz) # for creating heatmap plots
library(GGally) #for visualizing multiple models
library(tidymodels) # for creating generic model formulas
library(ggeffects)
library(performance) # for R-squared
library(pROC)#for ROC curve
library(scales)
library(patchwork)
library(correlation)
library(corrplot)


```

# Read in the tidy dataset

```{r Reading_Dataset}

here::i_am("CATAv2/03_TD_Graphs.Rmd")
df <- readRDS(here("CATAv2","R_ob","Cata_W3_sub_tidy.RDS" ))

dim(df)
# [1] 6216   31
#View(df)
#names(df)


```

# Assigning Labels to short names

```{r Assigning_labels }

# Loading named vector
source(here("CATAv2/Varnames.R"))

# Creating a named vector
data <- Hmisc::upData(df,labels = var.labels)

#View(data)

attributes(data$cq3g2a)

# changing labels to column names
names(data) <- colnames(label_to_colnames(data))  # uses sjlabelled



#Removing 3 work related variables
data$part_time_job <- NULL
data$work_for_business <- NULL
data$work_exp <- NULL
data$household_region <- NULL



# Looking the names
names(data)
```

# Exploratory plots

```{r corrplot}

data1 <- data %>% dplyr::mutate(across(high_inc:other, ~ factor(.x))) %>% dplyr::select(-ID, -schoolID) 

data1$school_type <- factor(data1$school_type)


responses <- names(data1)[1:10]
preds <- names(data1)[-(1:11)]



library(correlation)
# library(corrplot)
datap <- na.omit(data1[,c(responses,preds)])

datap %>% count(PC1_edu)

datap <- 
  datap %>% mutate(PC1_edu = fct_recode(PC1_edu, 
                                  School ="Prim",
                                  School="JC", 
                                  LC = "LC", 
                                  LC = "No_degree",
                                  Deg = "Deg",
                                  Deg = "Postgrad"))
datap %>% count(PC1_emp)

datap <- 
  datap %>% mutate(PC1_emp = fct_recode(PC1_emp, 
                                  Other ="Education",
                                  Work="Work", 
                                  Other = "Unemployed", 
                                  Other = "Home", 
                                  Other = "Other"))

for (i in 1:ncol(datap)) {
  f <- datap[[i]]
  if (class(f)[1] =="labelled")
    class(datap[[i]]) <- class(f)[-1]
}

datap$gender <- as.numeric(datap$gender)-1
names(datap)[11]<- "genderF"


datap$TY <- as.numeric(datap$TY =="Yes")
names(datap)[18]<- "TY-Y"

datap$PC1_emp <- as.numeric(datap$PC1_emp =="Work")
names(datap)[13]<- "PC1_empL_Work"

datap$school_type <- as.numeric(datap$school_type =="single")
names(datap)[names(datap)=="school_type"]<- "School_single"

datap$left_school <- as.numeric(datap$left_school =="Yes")
names(datap)[20]<- "Left_SchooL_Y"


for (i in 1:10) datap[[i]]<- as.numeric(datap[[i]])


pcorc <- correlation(datap, include_factors = T)
pcor <- as.data.frame(pcorc[,1:3])

varlevelsY <- rownames(as.matrix(pcorc))[1:10]
oY <- corrMatOrder(as.matrix(pcorc)[1:10,1:10], order="hclust", hclust.method="average")

varlevelsX <- rownames(as.matrix(pcorc))[-(1:10)]

cX <- as.matrix(pcorc)[-(1:10),-(1:10)]
# cX[2:7,2:7]<- 1
# cX[8:12,8:12]<- 1
oX <- corrMatOrder(cX, order="hclust", hclust.method="average")

newlevels <- c(varlevelsY[oY], varlevelsX[oX])

w <- which(unlist(map(strsplit(pcor$Parameter1, ".", fixed=T),1)) == 
             unlist(map(strsplit(pcor$Parameter2, ".", fixed=T),1)))
pcor[w,3]<- 0  # zero our correlations between levels of factor

ggplot(data=pcor, aes(x=Parameter1, y=Parameter2, fill=r))+ geom_tile()+
  geom_tile(aes(x=Parameter2, y=Parameter1))+
  xlab("")+ylab("")+
  scale_x_discrete(position = "top", limits=newlevels) + # x labels at top
  scale_y_discrete(limits = rev(newlevels))+ # reverse y axis
  scale_fill_gradient2(low ="red3", mid="white", high = "blue3", midpoint=0, limits=c(-1,1))+
  theme_light() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        aspect.ratio = 1 ,# for squares
        axis.text.x = element_text(angle = 30, hjust = 0)
  )


 
#
```



#modelling


```{r Main_effects}
# Making a vector for response variables
y <- data[,2:11] # omit other
names(y)

#get rid of IDs and create a dataframe for predictors
x <- data[,-c(2:12,which(names(data)=="ID"), which(names(data)== "schoolID"))]
names(x)
```


Rescaling all the numerical variables between 0 and 1

```{r}
sapply(x, class)
num_vars <- names(which(sapply(x, is.numeric)))

x$equalised_inc <- as.numeric(x$equalised_inc)
x$work_imp <- as.numeric(x$work_imp)
x$religion_imp <- as.numeric(x$religion_imp)
x$own_family_imp <- as.numeric(x$own_family_imp)
x$conscientious_score <- as.numeric(x$conscientious_score)
x$cognitive_scores <- as.numeric(x$cognitive_scores)
x$parental_interactions <- as.numeric(x$parental_interactions)
x$career_guidance <- as.numeric(x$career_guidance)

summary(x[,num_vars])
x[,num_vars] <- lapply(x[,num_vars],function(x) rescale(x,c(0,1)))
summary(x[,num_vars])

```




## Selecting interaction terms.
Selecting highly significant interaction (only dark blue)
```{r final_mod_without_order}


# Getting rid of some predictors
x1 <- x %>%
  dplyr::select(-c("TY",  "PC1_emp" , "PC1_edu", "left_school"))



# Refit model
fits.1 <- map(y, ~ glm(.x ~ . +own_family_imp*gender+
                           conscientious_score*gender+
                           school_type*gender ,
                       family="binomial",data=x1))



d_new <- purrr::map_dfr(fits.1, ~ broom::tidy(car::Anova(.x)), .id="response")
d1_new <- dplyr::filter(d_new, term != "Residuals")

cols <- c("blue", "cyan", "grey95")
final.plot <- modelHeatmap(d1_new, "term", "response", "p.value", xorder="increasing", yorder="increasing")+
  ggplot2::scale_fill_manual(values = cols)+ggplot2::xlab("Variables driving Work-Values") + ggplot2::ylab("Work-Values") + ggplot2::ggtitle("Significant Interactions with Gender")

final.plot
########################################################

```

```{r final_mod_ordered}
# plot the interactions and their main effects together
reorder_pred <- c("cognitive_scores","gender","work_imp","own_family_imp","gender:own_family_imp","conscientious_score","gender:conscientious_score","equalised_inc","career_guidance","religion_imp","household_region","school_type","gender:school_type","parental_interactions")



plot1 <- modelHeatmap(d1_new, "term", "response", "p.value", xorder=reorder_pred, yorder="increasing") +
  ggplot2::scale_fill_manual(values = cols) +
  ggplot2::xlab("Variables driving Work-Values") +
  ggplot2::ylab("Work-Values") +
  ggplot2::ggtitle("Final Model revealing potential Work Value Drivers")

plot1
```



# Coefficient Plots for all Work Values


```{r finl_model_coeff ,fig.width=14, fig.height=8}

#coefficients plot of main effects which donot have interactions
include_vars <- base::setdiff(names(x1),c("own_family_imp","gender","conscientious_score","school_type"))
```


```{r Cf_help ,fig.width=14, fig.height=8}
#plotting helping society

GGally::ggcoef_model(fits.1$help_society,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Helping Society")
```


```{r Cf_job ,fig.width=14, fig.height=8}
#plotting job security
ggcoef_model(fits.1$job_security,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Job Security")
```


```{r Cf_travel ,fig.width=14, fig.height=8}
#plotting travel abroad
ggcoef_model(fits.1$travel_abroad,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Travel Abroad")
```


```{r Cf_int ,fig.width=14, fig.height=8}
#plotting interesting job
ggcoef_model(fits.1$interesting_job,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Interesting Job")
```


```{r Cf_boss ,fig.width=14, fig.height=8}
#plotting being your own boss
ggcoef_model(fits.1$being_boss,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Being your own boss")
```


```{r Cf_flex ,fig.width=14, fig.height=8}
#plotting flexible hours
ggcoef_model(fits.1$flex_hours,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Flexible Hours")
```


```{r Cf_time ,fig.width=14, fig.height=8}
#plotting time off
ggcoef_model(fits.1$time_off,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Time Off")
```


```{r Cf_train ,fig.width=14, fig.height=8}
#plotting training opportunities
ggcoef_model(fits.1$training_opp,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Training Opportunity")
```


```{r Cf_high ,fig.width=14, fig.height=8}
#plotting high income
ggcoef_model(fits.1$high_inc,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for High Income")
```


```{r Cf_good ,fig.width=14, fig.height=8}
#plotting good for career
ggcoef_model(fits.1$good_for_career,include = include_vars, shape_values= c(19,1), point_stroke=.5, errorbar_coloured=F, point_size=3 )+ ggtitle("Main effects for Good for Career")



```

#Looking at the effects of variables

```{r ep-job ,fig.width=14, fig.height=8}
# Job Security with school type and gender
g1 <- ggpredict(fits.1$job_security, c("school_type", "gender"))
plot(g1)+geom_line(aes(linetype=group), 
            position=position_dodge(width=.25))
```


```{r ep-travelGender ,fig.width=14, fig.height=8}
# Travel Abroad with Conscientious score and gender
g2 <- ggpredict(fits.1$travel_abroad, c("conscientious_score", "gender"))
plot(g2)
```


```{r  ep-InterestingJob ,fig.width=14, fig.height=8}
# Interesting Job with Conscientious score and gender
g3 <- ggpredict(fits.1$interesting_job, c("conscientious_score", "gender"))
plot(g3)
```

# Effects plot of own family importance and gender for different work values
```{r ep-help ,fig.width=14, fig.height=8}
#plotting helping society
g4 <- ggpredict(fits.1$help_society, c("own_family_imp", "gender"))
plot(g4)
```


```{r ep-Jobsec ,fig.width=14, fig.height=8}
#plotting job security
g5 <- ggpredict(fits.1$job_security, c("own_family_imp", "gender"))
plot(g5)
```


```{r ep-travel ,fig.width=14, fig.height=8}
#plotting travel abroad
g6 <- ggpredict(fits.1$travel_abroad, c("own_family_imp", "gender"))
plot(g6)
```


```{r ep-Intjob ,fig.width=14, fig.height=8}
#plotting interesting job
g7 <- ggpredict(fits.1$interesting_job, c("own_family_imp", "gender"))
plot(g7)
```


```{r ep-boss ,fig.width=14, fig.height=8}
#plotting being your own boss
g8 <- ggpredict(fits.1$being_boss, c("own_family_imp", "gender"))
plot(g8)
```


```{r ep-time ,fig.width=14, fig.height=8}
#plotting time off
g9 <- ggpredict(fits.1$time_off, c("own_family_imp", "gender"))
plot(g9)
```


```{r ep-train ,fig.width=14, fig.height=8}
#plotting training opportunities
g10 <- ggpredict(fits.1$training_opp, c("own_family_imp", "gender"))
plot(g10)
```


```{r ep-high ,fig.width=14, fig.height=8}
#plotting high income
g11 <- ggpredict(fits.1$high_inc, c("own_family_imp", "gender"))
plot(g11)
```


```{r ep-goodCareer ,fig.width=14, fig.height=8}
#plotting good for career
g12 <- ggpredict(fits.1$good_for_career, c("own_family_imp", "gender"))
plot(g12)
```


```{r ep-flexhours ,fig.width=14, fig.height=8}
#plotting for flexible hours
g13 <- ggpredict(fits.1$flex_hours, c("own_family_imp", "gender"))
plot(g13)

    


```

