---
title: "02_DataSetup"
author: "Bharvi Dhall"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=5, figs="figsBD/")
```

# Setup CATA data by subsetting data from Wave 4
File contains the code to get a subset of data set (GUI wave 4) concerning career choices from GUI wave 4

# Set file path

```{r}
library(here)
here()
  
```

# Loading Libraries

```{r}
library(tidyverse)
library(labelled)
library(haven)
library(ggplot2)
library(here)
library(sjlabelled) 
library(Hmisc) #for assigning labels
library(manyModelViz) # for creating heatmap plots
library(GGally) #for visualizing multiple models
library(tidymodels) # for creating generic model formulas
library(ggeffects)
library(performance) # for R-squared
library(pROC)#for ROC curve
library(scales)
```


# Reading GUI dataset from wave 4

```{r}
here::i_am("GUI-Wave4-Analysis/01_ReadRawData.Rmd")
wave.4 <- readRDS(here("GUI-Wave4-Analysis","Robject", "Wave3_GUI_21yr.RDS"))

data.4 <- wave.4 %>%
  mutate_if(
    is.labelled,
    ~as_factor(.x)
  )

dim(data.4)

table(data.4$p2sexW4)

#View(data.4)
```




# Variables of Interest


cq4J5a - cq4J5l

Table 2: The table contains description of work values


| Name         | Value |   
|--------------|----------------------------|
| cq4J5a | a high income | E
| cq4J5b | offered good training opportunities | I
| cq4J5c | offered good promotion opportunities |
| cq4J5d | an interesting job | I
| cq4J5e | flexible working hours | E
| cq4J5f | generous holidays/time off |  E
| cq4J5g | a good step on the career ladder |
| cq4J5h | be your own boss |  
| cq4J5i | a job which allows you to be creative |  I
| cq4J5j | a job that is useful to society or helps other people|  I
| cq4J5k | job security |  E
| cq4J5l | opportunity to travel/work abroad |  E




Correlation plot of ranked choices
```{r}
wave_4_dict <- labelled::generate_dictionary(data.4)


wave_4_dict %>% filter(str_detect(variable, "cq4J5"))%>% select(pos, variable, label,col_type)

#For the purpose of this analysis, all the NA's have been coded as No.

#Taking subset of the variables of Interest
sub4 <- data.4[,c(1,504:515)]


# Loading named vector
source(here("GUI-Wave4-Analysis/vars.R"))

# Creating a named vector
dd <- Hmisc::upData(sub4,labels = var.labels)

#View(data)

attributes(dd$cq4J5a)

# changing labels to column names
names(dd) <- colnames(label_to_colnames(dd))  # uses sjlabelled

names(dd)

View(dd)

#correlation matrix

library("dplyr")
 dd %>%
   mutate(dplyr::across(high_inc:travel_abroad, as.numeric)) -> dd


# Correlation matrix of items
cormat <- dd %>%
  select(-1) %>%
  cor(method = "spearman", use = "pairwise.complete.obs")

# Activate the corrplot package
library("corrplot")

# Correlation matrix plot
corrplot(cormat, # correlation matrix
         order = "hclust", # hierarchical clustering of correlation
         tl.col = "black") # number of rectangles to draw around clusters

# males

summarise(dd)



#females



```


```{r}
#View(sub4) Do very imp as 1 else 0

#df4 <- sub4 %>%  mutate(across(starts_with("cq4J5"), ~ifelse(.x == 10,1,0)), across(starts_with("cq4J5"), ~replace_na(.x, 0)))

#View(df4)
#dim(df4)


# As per Catherine's suggestion for choosing top 3 choices
#changing work val to numeric
sub4_new <- sub4 %>%
  mutate(dplyr::across(starts_with("cq4J5"),as.character)) %>%
  mutate(dplyr::across(starts_with("cq4J5"),as.numeric)) 
#5190 X13

#getting rid of 999
sub4_new <- sub4_new %>%
  filter(cq4J5a != 99)
#5183 X13

# Select top 3 choices by first sorting by rank and then choices ranked at pos 10 or higher as there are 12 work values to choose from and taking them as one
df4 <- sub4_new %>%
  select(-1) %>%
  mutate(across(.cols=tidyselect::everything(), ~replace_na(.x, 0))) %>%
  apply(1, function(x) ifelse(x >= sort(x)[10],1,0))  %>% t() %>% as.data.frame() %>% add_column(ID=sub4_new$ID,.before = 1)

rownames(df4) <- NULL
# Loading named vector
source(here("GUI-Wave4-Analysis/vars.R"))

# Creating a named vector
data <- Hmisc::upData(df4,labels = var.labels)

#View(data)

attributes(data$cq4J5a)

# changing labels to column names
names(data) <- colnames(label_to_colnames(data))  # uses sjlabelled

names(data)

# ratings <- sample(10, 13, replace=T)
# ratings
# sort(ratings)
# ifelse(ratings >= sort(ratings)[11], 1,0)

```





# Merge

```{r}



#Load R object of Wave 3
wave.3 <- readRDS(here("GUI-Wave4-Analysis","Robject", "Cata_W3_sub_tidy.RDS"))


w4_sub <- merge(wave.3,data,by.x = "ID", by.y = "ID", all= FALSE)
# after merging only 4811 left

dim(w4_sub)

#View(w4_sub)
```

## Fit logistic regression with main effects

We are fitting Logistic regression using all predictors, and then using car::Anova to find the effect of each predictor on the response variable (Is there any significant association between the drivers and workvalues?)


```{r Main_effects}
# Making a vector for response variables
y <- w4_sub[,32:43] # omit other
names(y)


#get rid of IDs and create a dataframe for predictors
x <- w4_sub[,-c(2:12,32:43,1,27)]

# Loading named vector
source(here("GUI-Wave4-Analysis/vars.R"))
# Creating a named vector
x <- Hmisc::upData(x,labels = var.labels)


# changing labels to column names
names(x) <- colnames(label_to_colnames(x))  # uses sjlabelled

names(x)

#Removing 3 work related variables
x$part_time_job <- NULL
x$work_for_business <- NULL
x$work_exp <- NULL
x$household_region <- NULL
```


```{r Main_effects2}
# Fitting logistic regression on main effects
fits <- map(y, ~ glm(.x ~ . ,family="binomial",data=x))
f <- purrr::map_dfr(fits, ~ broom::tidy(car::Anova(.x)), .id="response")
f1 <- dplyr::filter(f, term != "Residuals")

# make heatmap
cols <- c("blue", "cyan", "grey95")
modelHeatmap(f1, "term", "response", "p.value", xorder="increasing", yorder="increasing")+
  ggplot2::scale_fill_manual(values = cols)+ ggplot2::xlab("Variables driving Work-Values") + ggplot2::ylab("Work-Values") + ggplot2::ggtitle("Heatmap of 10 Logistic Regression fits of the Work-Values")

# cq3g4f(work imp)  coming significant in main effects
```


## Fit the logistic regressions and check for interactions with gender and fit model with main effects and significant interactions.

```{r}

fits.2 <- map(y, ~ glm(.x ~ (.-gender)*gender ,family="binomial",data=x))
d <- purrr::map_dfr(fits.2, ~ broom::tidy(car::Anova(.x)), .id="response")
d1 <- dplyr::filter(d, term != "Residuals")
#y <- "response"
d1 <- d1[grep("gender:|:gender", d1$term),] # to check whether term has "gender:"
d1$term <- sub("gender:|:gender", "",d1$term) # to replace "gender:" with "" for values of terms for all interactions

cols <- c("blue", "cyan", "grey95")
modelHeatmap(d1, "term", "response", "p.value", xorder="increasing", yorder="increasing")+
  ggplot2::scale_fill_manual(values = cols)+ ggplot2::xlab("Variables associated with Work-Values") + ggplot2::ylab("Work-Values") + ggplot2::ggtitle("Interactions with Gender")
```

The significant interactions with gender are: work_imp:gender, (highly) TY:gender (sig)

- get rid of vars: PC1_emp and left school


## Selecting interaction terms.
Selecting highly significant interaction (only dark blue)
```{r final_model}
sig_int_vars2 <- d1 %>% filter(p.value<=0.01) %>% select(term) %>% unique() %>% deframe()

# Getting rid of some predictors
x1 <- x %>%
  dplyr::select(-c("PC1_emp" , "left_school"))



# Refit model
fits.1 <- map(y, ~ glm(.x ~ . +work_imp*gender+
                           TY*gender ,
                       family="binomial",data=x1))



d_new <- purrr::map_dfr(fits.1, ~ broom::tidy(car::Anova(.x)), .id="response")
d1_new <- dplyr::filter(d_new, term != "Residuals")

cols <- c("blue", "cyan", "grey95")
final.plot <- modelHeatmap(d1_new, "term", "response", "p.value", xorder="increasing", yorder="increasing")+
  ggplot2::scale_fill_manual(values = cols)+ggplot2::xlab("Variables associated with Work-Values") + ggplot2::ylab("Work-Values") + ggplot2::ggtitle("Significant Interactions with Gender")

final.plot
########################################################
# plot the interactions and their main effects together
reorder_pred <- c("cognitive_scores","gender","conscientious_score","TY","gender:TY","work_imp","gender:work_imp","equalised_inc","PC1_edu","career_guidance","own_family_imp","parental_interactions","religion_imp","school_type")



plot1 <- modelHeatmap(d1_new, "term", "response", "p.value", xorder=reorder_pred, yorder="increasing") +
  ggplot2::scale_fill_manual(values = cols) +
  ggplot2::xlab("Potential Factors") +
  ggplot2::ylab("Work-Values") +
  ggplot2::ggtitle("Final Model revealing Variables associated with Work Values")

plot1


```



# Looking at interaction between gender and TY for gender for good for career, promotion opp, training opp

```{r ep-job ,fig.width=14, fig.height=8}
# Job Security with school type and gender
g1 <- ggpredict(fits.1$good_for_career, c("TY", "gender"))
plot(g1)+geom_line(aes(linetype=group), 
            position=position_dodge(width=.25))
```

```{r ep-job ,fig.width=14, fig.height=8}
# Job Security with school type and gender
g2 <- ggpredict(fits.1$promotion_opp, c("TY", "gender"))
plot(g2)+geom_line(aes(linetype=group), 
            position=position_dodge(width=.25))
```

```{r ep-job ,fig.width=14, fig.height=8}
# Job Security with school type and gender
g3 <- ggpredict(fits.1$training_opp, c("TY", "gender"))
plot(g3)+geom_line(aes(linetype=group), 
            position=position_dodge(width=.25))
```

# Some exploratory plot

```{r}

library(ggupset)
library(here)
library(tidyverse)
library(patchwork)



df1<- as.matrix(data)

catavars <- c("high_inc", "training_opp","promotion_opp","interesting_job","flex_hours","time_off", "good_for_career", "being_boss" ,  "creative_job","help_society","job_security" ,"travel_abroad")
df4$choice <- lapply(1:nrow(df1), function(i) catavars[df1[i,]==1])



#df3 <- dim(df4[sapply(df4$choice, length) ==9,])

#choose 3 - 1438
# choose 4- 1156
#choose 10 -97

names(df4)

saveRDS( df4, here("GUI-Wave4-Analysis","Robject", "df4.RDS"))
```


## Basic bar charts

```{r, fig.width=7}
names(df4)[2:13]<- catavars

#adding gender
dfgen <- data.4 %>%
  filter(cq4J5a != "99")

df4$gender <- dfgen$p2sexW4

#recode gender
df4  <- 
  df4 %>% mutate(gender = fct_recode(gender, Male = "1",
                                       Female ="2"))
  
colnames(df4)

df4 %>% pivot_longer(cols=high_inc:travel_abroad, names_to = "selection")  %>% 
  filter(value==1)  %>% 
  count(selection,gender) %>% 
  ggplot(aes(x=fct_reorder(selection,n, .desc=T), y=n)) + 
  geom_col(aes(fill=gender), position="dodge")+   xlab("Selection") + ylab("Count")+theme(axis.text=element_text(angle = 60,size=10, vjust = 0.5, hjust =1),
        axis.title=element_text(size=14,face="bold"))


names(df4)
```


# Plot extrinsic choices and extrinsic

```{r}
col.ext <- c("high_inc", "flex_hours", "time_off", "job_security", "travel_abroad", "promotion_opp")

df4 %>% pivot_longer(cols=col.ext, names_to = "selection")  %>% 
  filter(value==1)  %>% 
  count(selection,gender) %>% 
  ggplot(aes(x=fct_reorder(selection,n, .desc=T), y=n)) + 
  geom_col(aes(fill=gender), position="dodge")+   xlab("Selection") + ylab("Count")+theme(axis.text=element_text(angle = 60,size=10, vjust = 0.5, hjust =1),
        axis.title=element_text(size=14,face="bold"))


col.int <- c("interesting_job", "training_opp", "creative_job", "help_society")

df4 %>% pivot_longer(cols=col.int, names_to = "selection")  %>% 
  filter(value==1)  %>% 
  count(selection,gender) %>% 
  ggplot(aes(x=fct_reorder(selection,n, .desc=T), y=n)) + 
  geom_col(aes(fill=gender), position="dodge")+   xlab("Selection") + ylab("Count")+theme(axis.text=element_text(angle = 60,size=10, vjust = 0.5, hjust =1),
        axis.title=element_text(size=14,face="bold"))

```

# plot choices = 9 very Impp

```{r}
 
dim(df4 %>% select_if(~max(.)>9))
```



There are about equal numbers of men and women in the data.
Women pick travel and helpSoc more than men, and men pick highInc and flexHrs more than women.

```{r, fig.width=7}
df4 %>% pivot_longer(cols=high_inc:travel_abroad, names_to = "selection")  %>% 
  filter(value==1)  %>% 
  count(selection,gender) %>% 
  ggplot(aes(x=fct_reorder(selection,n, .desc=T,.fun=diff), y=n)) + 
  geom_line(color="grey50", aes(group=selection))+
  geom_point(aes(color=gender), size=4)+ 
  xlab("Selection") + ylab("Count")+theme(axis.text=element_text(angle = 60,size=10, vjust = 0.5, hjust =1),
        axis.title=element_text(size=14,face="bold"))
```

## Basic upset plot, showing top 10 choices
```{r}
df3 %>% ggplot(aes(x=choice)) +
  geom_bar() +
  scale_x_upset(n_intersections = 10)
```

About 10% of respondents picked the first choice. These are also the most popular individual choices.
There are 165 possible choices and about half the respondents chose these 10.
All of the choices involve interesting job. And most involve highInc.
Notice some of the choices (trainingOp, boss, other) do not appear here.

We can check how these top choices are split between males and females

## Comparing males and females

```{r}
df3 %>% ggplot(aes(x=choice, fill=gender)) +
  geom_bar(position="fill") +
  scale_x_upset(n_intersections = 10)

#Orange males, blue females
```
