---
title: "03_EDA"
author: "Bharvi Dhall"
date: "2022-10-16"
output: html_document
---

# Setup CATA data by subsetting data from Wave 4
File contains the code to get a subset of data set (GUI wave 4) concerning career choices from GUI wave 4

# Set file path

```{r}
library(here)
here()
  
```

# Loading Libraries

```{r}
library(tidyverse)
library(labelled)
library(haven)
library(ggplot2)
library(here)
library(sjlabelled) 
library(Hmisc) #for assigning labels
library(manyModelViz) # for creating heatmap plots
library(GGally) #for visualizing multiple models
library(tidymodels) # for creating generic model formulas
library(ggeffects)
library(performance) # for R-squared
library(pROC)#for ROC curve
library(scales)
```


# Reading GUI dataset from wave 4

```{r}
here::i_am("GUI-Wave4-Analysis/01_ReadRawData.Rmd")
wave.4 <- readRDS(here("GUI-Wave4-Analysis","Robject", "Wave3_GUI_21yr.RDS"))

data.4 <- wave.4 %>%
  mutate_if(
    is.labelled,
    ~as_factor(.x)
  )

dim(data.4)

table(data.4$p2sexW4)

#View(data.4)
```




# Variables of Interest


cq4J5a - cq4J5l

Table 2: The table contains description of work values


| Name         | Value |   
|--------------|----------------------------|
| cq4J5a | a high income | E
| cq4J5b | offered good training opportunities | I
| cq4J5c | offered good promotion opportunities |
| cq4J5d | an interesting job | I
| cq4J5e | flexible working hours | E
| cq4J5f | generous holidays/time off |  E
| cq4J5g | a good step on the career ladder |
| cq4J5h | be your own boss |  
| cq4J5i | a job which allows you to be creative |  I
| cq4J5j | a job that is useful to society or helps other people|  I
| cq4J5k | job security |  E
| cq4J5l | opportunity to travel/work abroad |  E




Correlation plot of ranked choices
```{r}
wave_4_dict <- labelled::generate_dictionary(data.4)


wave_4_dict %>% filter(str_detect(variable, "cq4J5"))%>% select(pos, variable, label,col_type)

#For the purpose of this analysis, all the NA's have been coded as No.

#Taking subset of the variables of Interest
sub4 <- data.4[,c(1,504:515)]

sub4 <- sub4 %>%
  mutate(dplyr::across(starts_with("cq4J5"),as.character)) %>%
  mutate(dplyr::across(starts_with("cq4J5"),as.numeric)) 


# Loading named vector
source(here("GUI-Wave4-Analysis/vars.R"))

# Creating a named vector
dd <- Hmisc::upData(sub4,labels = var.labels)

#View(data)

attributes(dd$cq4J5a)

# changing labels to column names
names(dd) <- colnames(label_to_colnames(dd))  # uses sjlabelled

names(dd)


```


#merge 

```{r}
# Merge

#Load R object of Wave 3
wave.3 <- readRDS(here("GUI-Wave4-Analysis","Robject", "Cata_W3_sub_tidy.RDS"))

View(wave.3)

df.merge <- merge(wave.3,dd,by.x = "ID", by.y = "ID", all= FALSE)
# after merging only 4811 left

dim(df.merge)

#View(df.merge)
```


```{r}
data_to_work<-df.merge[,c(1,13,32:43)]
#View(data_to_work)

saveRDS(data_to_work, here("GUI-Wave4-Analysis","Robject", "dw.RDS"))



dat<-data_to_work %>%
  pivot_longer(cols = 3:14, names_to = c("work_value"), values_to = "score") %>%
  na.omit %>%
  group_by(p2sexW3, work_value, score) %>%
  summarise(freq = n())

View(dat)


r<-dat %>%
  ggplot(
    aes(
      x      = work_value, 
      y      = score +  (freq / max(freq) * .95) / 2, 
      height = (freq / max(freq) * .95), 
      width  = .95,
      fill   = as.factor(score),
      label  = freq
      )
    ) +
  geom_tile(show.legend = FALSE) +
  facet_wrap(~ p2sexW3) +
  labs(x = 'Work Values', y = 'Importance') +
  scale_y_discrete(limits = c("0","1","2","3","4","5","6","7","8","9","10")) +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.y  = element_text(hjust = 0, vjust = -0.025),
    axis.ticks.y = element_blank()
  )+
  theme_bw()
r+theme(axis.text.x  = element_text(angle = 30,size = 10, hjust = 1, vjust = 0.9))






p<-dat %>%
  ggplot(
    aes(
      x      = work_value, 
      y      = score +  (freq / max(freq) * .95) / 2, 
      height = (freq / max(freq) * .95), 
      width  = .95,
      fill   = p2sexW3,
      label  = freq
      )
    ) +
  geom_tile(show.legend = FALSE,position = "dodge") +
  labs(x = 'Work Values', y = 'Importance') +
  scale_y_discrete(limits = c("0","1","2","3","4","5","6","7","8","9","10")) +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.y  = element_text(hjust = 0, vjust = -0.025),
    axis.ticks.y = element_blank()
  )+
  theme_bw()

p + theme(axis.text.x  = element_text(angle = 30,size = 10, hjust = 1, vjust = 0.9))
```


